Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-06-15 paper.qmd

Tools for estimating geographical distribution of cycling potential have advanced substantially in recent years.
Developed by researchers and developers in academic, public and private sectors, they have evolved from simple area-based static models tailored to specific regions to more complex and (in some cases) more generalizable tools.
These tools can be considered as a specific application of the concept of planning support systems (PSS) , which are designed to help planners and decision-makers to make more informed decisions [@geertman2009].
Outputs include maps with results available at several levels of analysis with levels of availability ranging from being only available to researchers and in static maps to open access mapping systems available to the public.

@@

```{r}
#| label: tbl-tools
# # For PDF:
# knitr::kable(tools, caption = "Summary of existing tools for estimating cycling potential.", booktabs = TRUE) |>
#   kableExtra::kable_styling(latex_options = "striped", full_width = FALSE) |>
#   kableExtra::column_spec(1, width = "13em") |>
#   kableExtra::column_spec(2, width = "15em") |>
#   kableExtra::column_spec(3, width = "12em") |>
#   kableExtra::column_spec(4, width = "5em")
# For HTML:
knitr::kable(tools, caption = "Summary of existing tools for estimating cycling potential.")
```

@@

Other tools that generate areal data to prioritize investment include a study that generated a 'prioritization index' for Montreal, Canada [@larsen_build_2013], a 'usage intensity index' based on stated preferences in Belo Horizonte, Brazil, and a Cycling Potential Tool (CPT) for Scotland [@phillips_development_2017].
The CPT is composed of the base environmental module — based on eight weighted factors (population density, hilliness, physical barriers, access to services, existing cycling mode share, distance to work and school, and road speed) — and the quality of service module, which evaluates each area of interest based on eight factors (surface condition, adjacent cyclists, comfort factor, conflict, distance between junctions, slope, access to services, and origin/destination).

A more recent areal-based approach is the designed the Gross Potential for Cycling (GPC) tool to prioritize areas for cycling infrastructure and other cycling measures [@silva_gross_2021].
The GPC uses two sets of indicators: population-based indicators (age, potential demand density, employment density and motorization rate) and area-based indicators (accessibility, education, public transport, connectivity, land use mix and relative performance).
These indicators are calculated considering topography, road hierarchy, and average congestion and presented on a scale from 5 to 10.
They are then combined into an overall score, weighted by their impact on cycling according to the literature.
The ranking of areas by the GPC are [publicly available](https://boost.up.pt/en/ferramentas/gpc) and its practical value has been assessed through workshops [@silva_revealing_2022].

Route-based tools emerged with increasing availability of origin-destination data, routing engines that can assign trips to networks, and improvements in computer hardware and software needed to generate route networks.
A prominent example is the the Propensity to Cycle Tool (PCT), first developed to estimate current and future levels of cycling at desire line, zone, route, and route network levels for case study cities [@lovelace_mapping_2016].
The approach was scaled-up to estimate estimate the potential benefits of uptake at zone and desire line levels nationally, and launched as a publicly available web application in 2017 [@lovelace2017].
Extensions of the PCT approach have included estimation of benefits at the individual level [@woodcock2018], addition of travel to school network [@goodman2019], and improved modeling of impacts on health, environmental and distributional outcomes [@woodcock2021].
Initially developed just for England, the PCT was extended to cover all of Wales (for commuter data only) in 2018.

@@

@@

```{r}
#| label: tbl-trip-purposes
#| tbl-cap: "Trip purposes used in the CRUSE Tool."
# For PDF:
# knitr::kable(trip_purposes, booktabs = TRUE) |>
#   kableExtra::kable_styling(latex_options = "striped", full_width = FALSE, font_size = 8) |>
#   kableExtra::column_spec(1, width = "5em") |>
#   kableExtra::column_spec(2, width = "12em") |>
#   kableExtra::column_spec(3, width = "15em") |>
#   kableExtra::column_spec(4, width = "15em")
# For HTML:
knitr::kable(trip_purposes, caption = "Trip purposes used in the CRUSE Tool.")
```

>>>>

Tools for estimating geographical distribution of cycling potential have advanced substantially in recent years.
Developed by researchers and developers in academic, public and private sectors, they have evolved from simple area-based static models tailored to specific regions to more complex and (in some cases) more generalizable tools.
These tools can be considered as a specific application of the concept of planning support systems (PSS), which are designed to help planners and decision-makers to make more informed decisions [@geertman2009].
Outputs include maps with results available at several levels of analysis with levels of availability ranging from being only available to researchers and in static maps to open access mapping systems available to the public.

@@

```{r}
#| label: tbl-tools
#| tbl-cap: "Summary of existing tools for estimating cycling potential."
# # For PDF:
# knitr::kable(tools, caption = "Summary of existing tools for estimating cycling potential.", booktabs = TRUE) |>
#   kableExtra::kable_styling(latex_options = "striped", full_width = FALSE) |>
#   kableExtra::column_spec(1, width = "13em") |>
#   kableExtra::column_spec(2, width = "15em") |>
#   kableExtra::column_spec(3, width = "12em") |>
#   kableExtra::column_spec(4, width = "5em")
# For HTML:
knitr::kable(tools)
```

@@

Other tools that generate areal data to prioritize investment include a study that generated a<span style="color: blue;"> 'bikeability index' for the Metro Vancouver region [@winters_mapping_2013] and later called Bike Score and available for over 160 US and Canadian cities [@winters_bike_2016] -- its methodology is [publicly available](https://www.walkscore.com/bike-score-methodology.shtml); a </span>'prioritization index' for Montreal, Canada [@larsen_build_2013]; a 'usage intensity index' based on stated preferences in Belo Horizonte, Brazil [@zhang_prioritizing_2014]; and a Cycling Potential Tool (CPT) for Scotland [@phillips_development_2017].
The CPT is composed of the base environmental module — based on eight weighted factors (population density, hilliness, physical barriers, access to services, existing cycling mode share, distance to work and school, and road speed) — and the quality of service module, which evaluates each area of interest based on eight factors (surface condition, adjacent cyclists, comfort factor, conflict, distance between junctions, slope, access to services, and origin/destination).

A more recent areal-based approach is the designed Gross Potential for Cycling (GPC) tool to prioritize areas for cycling infrastructure and other cycling measures  [@silva_gross_2021].
The GPC uses two sets of indicators: population-based indicators (age, potential demand density, employment density and motorization rate) and area-based indicators (accessibility, education, public transport, connectivity, land use mix and relative performance).
These indicators are calculated considering topography, road hierarchy, and average congestion and presented on a scale from 5 to 10.
They are then combined into an overall score, weighted by their impact on cycling according to the literature.
The ranking of areas by the GPC are [publicly available](https://boost.up.pt/en/ferramentas/gpc) and its practical value has been assessed through workshops [@silva_revealing_2022].
<span style="color: blue;">The GPC tool has also been used to investigate the combination of factors contributing to high levels cycling potential [@lopes_impact_2021]. 
Another recent areal-based approach is the Relative Equity Impact of Bicycle Planning (TIRE) [@cunha_assessing_2023]. 
The TIRE assesses the impact of cycling network allocation on the accessibility levels of different socioeconomic groups.</span>

Route-based tools emerged with increasing availability of origin-destination data, routing engines that can assign trips to networks, and improvements in computer hardware and software needed to generate route networks.
A prominent example is the the Propensity to Cycle Tool (PCT), first developed to estimate current and future levels of cycling at desire line, zone, route, and route network levels for case study cities [@lovelace2016].
The approach was scaled-up to estimate estimate the potential benefits of uptake at zone and desire line levels nationally, and launched as a publicly available web application in 2017 [@lovelace2017].
Extensions of the PCT approach have included estimation of benefits at the individual level [@woodcock2018], addition of travel to school network [@goodman2019], and improved modeling of impacts on health, environmental and distributional outcomes [@woodcock2021].
Initially developed just for England, the PCT was extended to cover all of Wales (for commuter data only) in 2018.

@@

<span style="color: blue;">Another route-based tool is The Cycling Routing Algorithm for Network Connectivity (CRANC) [@gehrke_cycling-focused_2020]. 
The CRANC tool evaluates the benefits of new bicycle facilities for the accessibility of various populations and neighbourhoods. </span>

@@

```{r}
#| label: tbl-trip-purposes
#| tbl-cap: "Trip purposes used in the CRUSE Tool."
# For PDF:
# knitr::kable(trip_purposes, booktabs = TRUE) |>
#   kableExtra::kable_styling(latex_options = "striped", full_width = FALSE, font_size = 8) |>
#   kableExtra::column_spec(1, width = "5em") |>
#   kableExtra::column_spec(2, width = "12em") |>
#   kableExtra::column_spec(3, width = "15em") |>
#   kableExtra::column_spec(4, width = "15em")
# For HTML:
knitr::kable(trip_purposes)
```
